---
title: "pCR NAC WGCNA"
author: "Javier Moreno Venegas"
date: "29/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up R session

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
workingDir = "/Users/javi/Desktop/R-workspace/chemoResNAC_wgcna"
setwd(workingDir)
set.seed(12345)
options(stringsAsFactors = FALSE)

# Load required libraries
source(paste(getwd(), "Code/0_loadLibraries.R", sep='/'))
loadpkg("NanoStringQCPro")
loadpkg("dplyr")
loadpkg("WGCNA")
loadpkg("biomaRt")
#loadpkg("iRefR")
loadpkg("clusterProfiler")
loadpkg("rebus")
loadpkg("igraph")
loadpkg("diffuStats")

allowWGCNAThreads() # multi-threading within WGCNA. RStudio. For r, Rscript use enableWGCNAThreads()
```

## Data input  

### Load RCC files 

```{r, echo=TRUE}
DataDir <- "Data" # ../../data
rccDir <- file.path(DataDir, "RCCData")

## Process file names: <sample>_<key1>_<key2>_<subtype>.RCC
## We have 204 files. There are 6 files without <subtype>
## Key 1 contains pre/post info. 0 = PRE 
pre_samples = data.frame(fname=list.files(rccDir, pattern="*.RCC")) %>%
  tidyr::separate(fname,c("sample","key1","key2","subtype"),sep="_",remove=F) %>%
  tidyr::separate(subtype, c("subtype", "ext"), sep="\\.") %>%
  dplyr::select (fname,sample, key1,key2) %>%
  dplyr::filter(key1 == 0)

## Files without subtype has key2 = 'key2.RCC' so we have to remove '.RCC' to avoid further errors
for (i in grep(pattern = '\\.' ,pre_samples$key2)) {
  pre_samples$key2[i] <- unlist(strsplit(pre_samples$key2[i], '\\.'))[1]
}

rccSet <- newRccSet(
  rccFiles               = file.path(rccDir, pre_samples$fname)
  #,rccCollectorToolExport = file.path(exampleDataDir, "nSolver", "RCC_collector_tool_export.csv")
  ,rlf                    = file.path(DataDir,"RLFData/NS_CancerPath_C2535.rlf") #rlf file in RLFData subdir
  #,cdrDesignData          = file.path(exampleDataDir, "CDR", "CDR-DesignData.csv")
  #,extraPdata             = file.path(exampleDataDir, "extraPdata", "SampleType.txt")
  #,blankLabel             = "blank"
  ,experimentData.name    = "PRE_pCR"
  ,experimentData.lab     = "LBMC"
  ,experimentData.contact = "javimoreno1806@uma.es"
  ,experimentData.title   = "WGCNA BC_NAC dataset PRE samples"
  ,experimentData.abstract= "Gene co-expression network to find genes associated with pCR in pre-NAC samples."
)
```  

### Expression data preprocessing and cleaning   

```{r, echo=TRUE}
## Preprocessing and normalization
norm_rccSet <- preprocRccSet(rccSet = rccSet, doPosCtrlNorm = T, doContentNorm =T, normMethod = "housekeeping",bgReference = "negative")

## QC report
qc_rccSet <- makeQCReport(norm_rccSet, "QC_report"
                          , outputDir = paste(getwd(),'Results',sep='/'))

## Expression matrix with normalised data
ncounts <- assayData(qc_rccSet)$normData

## Apply QC filter

keepFiles <- dplyr::filter(qc_rccSet@phenoData@data, TechnicalFlags == FALSE)$FileName
ncounts <- ncounts[, keepFiles]

## Remove control genes (positive, negative, housekeeping)
rm_control <- (qc_rccSet@featureData@data %>% 
                 tibble::rownames_to_column(var = "FullName") %>%
                 dplyr::select(FullName, CodeClass) %>%
                 dplyr::filter(!grepl("Pos",CodeClass)) %>%
                 dplyr::filter(!grepl("Neg", CodeClass)) %>%
                 dplyr::filter(!grepl("Spike", CodeClass)) %>%
                 dplyr::filter(!grepl("Ligati", CodeClass)) %>%
                 dplyr::filter(!grepl("Housekee", CodeClass)))$FullName

ncounts <- ncounts[rm_control, ]

## Change columns from file name to sample name
pre_samples <- filter(pre_samples, fname %in% colnames(ncounts))
colnames(ncounts) <- pre_samples$sample

## Transpose the expression data
datExpr0 = as.data.frame(t(ncounts))

## Check samples for missing expression values
gsg = goodSamplesGenes(datExpr0, verbose = 3);
gsg$allOK

if (!gsg$allOK){
  # Optionally, print the gene and sample names that were removed:
  if (sum(!gsg$goodGenes)>0)
    printFlush(paste("Removing genes:", paste(names(datExpr0)[!gsg$goodGenes], collapse = ", ")));
  if (sum(!gsg$goodSamples)>0)
    printFlush(paste("Removing samples:", paste(rownames(datExpr0)[!gsg$goodSamples], collapse = ", ")));
  # Remove the offending genes and samples from the data:
  datExpr0 = datExpr0[gsg$goodSamples, gsg$goodGenes]
}

## Clustering to detect outliers
sampleTree = hclust(dist(datExpr0), method = "average");

## Plot the sample tree: Open a graphic output window of size 12 by 9 inches
## The user should change the dimensions if the window is too large or too small.

sizeGrWindow(12,9)   
# pdf(file = "sampleClustering.pdf", width = 12, height = 9);
par(cex = 0.6);
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5,
     cex.axis = 1.5, cex.main = 2)

# Plot a line to show the cut
abline(h = 65, col = "red");
# dev.off()
#################################

## Determine cluster under the line
clust = cutreeStatic(sampleTree, cutHeight = 65, minSize = 10)
table(clust)

keepSamples = (clust==1)
datExpr = datExpr0[keepSamples,]
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)
```  


We  keep `r nsamples` samples after QC and clustering.

### Load clinical data  

```{r, echo=TRUE}
pCRDat <- pre_samples %>% dplyr::select(sample,key2)

names(pCRDat) <- c("sample", "pCR")
pCRDat = pCRDat[keepSamples,]

pCR.to.chemores <- function(pcr){
  if(pcr == 0){return(1)}
  if(pcr == 1){return(0)}
}

chemores <- pCRDat
chemores$ChemoRes<- sapply(chemores$pCR, pCR.to.chemores)
chemores <- dplyr::select(chemores, sample, ChemoRes)
chemores$ChemoRes <- as.numeric(chemores$ChemoRes)
datTrait <- dplyr::select(chemores, ChemoRes)
rownames(datTrait) <- chemores$sample
```  

### Plot a cluster of the samples and the pCR values as color  

```{r}
# White means low (pCR =0) and red means pCR =1

sampleTree2 = hclust(dist(datExpr), method = "average")
# Convert traits to a color representation: white means low, red means high, grey means missing entry
traitColors = numbers2colors(datTrait, signed = FALSE);
# Plot the sample dendrogram and the colors underneath.
# pdf("dendogram.pdf",  width = 16, height = 9)
plotDendroAndColors(sampleTree2, traitColors,
                    groupLabels = names(datTrait),
                    main = "Sample dendrogram chemoresistance heatmap")
# dev.off()
save(datExpr, datTrait, file = "Data/RData/chemoresNAC_wgcna-01-dataInput.RData")
```
  
## Network construction and module detection  

### Soft-thresholding power selection  

```{r, echo=TRUE}
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to=20, by=2));
# Call the network topology analysis function
sft = pickSoftThreshold(datExpr, powerVector = powers, verbose = 5);

# Plot the results:
sizeGrWindow(9, 5);
par(mfrow = c(1,2));
#pdf("softpower_threshold.pdf",  width = 9, height = 9)
cex1 = 0.9;

# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red");

# This line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")

# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
abline(h=0,col='red')
#dev.off()
```  

### Build a Topological Overlaping Matrix from adjacency matrix  
```{r message=FALSE, include=FALSE}
softPower = 6
adjacency = adjacency(datExpr, power = softPower);

# Turn adjacency into topological overlap
TOM = TOMsimilarity(adjacency);
dissTOM = 1-TOM
```  

### Gene clustering on TOM-based dissimilarity  
```{r}
# Call the hierarchical clustering function
geneTree = hclust(as.dist(dissTOM), method = "average");

# Plot the resulting clustering tree (dendrogram)
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE, hang = 0.04)
```  

### Module identification  
```{r message=FALSE}
# Since I get few modules, I'll try with smaller modules:
minModuleSize = 20;
# Module identification using dynamic tree cut:
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM,
                            deepSplit = 2, pamRespectsDendro = FALSE,
                            minClusterSize = minModuleSize);
table(dynamicMods)
```  

### Gene clustering modules  
```{r}
# Convert numeric labels into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
# Plot the dendrogram and colors underneath
sizeGrWindow(12,9)
# pdf("gene_clustering_modules.pdf",  width = 12, height = 9)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene dendrogram and module colors")
```  

### Modules clustering  

The Dynamic Tree Cut may identify modules whose expression profiles are very similar. It may be prudent to merge such modules since their genes are highly co-expressed. To quantify co-expression similarity of entire modules, we calculate their eigengenes and cluster them on their correlation:  

```{r}
# Calculate eigengenes
MEList = moduleEigengenes(datExpr, colors = dynamicColors)
MEs = MEList$eigengenes
# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs);
# Cluster module eigengenes
METree = hclust(as.dist(MEDiss), method = "average");
# Plot the result
sizeGrWindow(7, 6)
#pdf("modules-merge.pdf",  width = 7, height = 6)
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")

MEDissThres = 0.25 # Corresponds with a correlation of 0.75
# Plot the cut line into the dendrogram
abline(h = MEDissThres, col = "red")
#dev.off()
```  

### Merge modules with highly co-expressed genes  
```{r echo=TRUE, message=FALSE, paged.print=TRUE}
# Call an automatic merging function
merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)
# The merged module colors
mergedColors = merge$colors;
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs;

sizeGrWindow(12, 9)
# pdf(file = "gene_clustering_modules-merge.pdf", wi = 12, he = 9)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)
# dev.off()
```  

### Rename, label and save  
```{r}
# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs = mergedMEs;
# Save module colors and labels for use in subsequent parts
save(MEs, moduleLabels, moduleColors, geneTree, file = "Data/RData/chemoresNAC_wgcna-02-networkConstruction.RData")
```  


  
## Relate modules to clinical chemoresistance  

```{r}
# Load saved data (datExpr and datTrait)
load("Data/RData/chemoresNAC_wgcna-01-dataInput.RData")
load("Data/RData/chemoresNAC_wgcna-02-networkConstruction.RData")
```  

### Quantifing module-trait associations  

In this analysis we would like to identify modules that are significantly associated with the measured clinical traits. Since we already have a summary profile (eigengene) for each module, we simply correlate eigengenes with external traits and look for the most significant associations:  

```{r}
# Define numbers of genes and samples
nGenes = ncol(datExpr);
nSamples = nrow(datExpr);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, datTrait, use = "p");
# moduleTraitCor = bicor(MEs, datTrait, use="p",robustY=FALSE, maxPOutliers=0.1)
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);

sizeGrWindow(12,6)
# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
                   signif(moduleTraitPvalue, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6.5, 8, 3, 3));
# Display the correlation values within a heatmap plot

#pdf(file = "Results/module-trait.pdf", wi = 8, he = 10)
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(datTrait),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 1.0,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))

#dev.off()
```  

Module eigengene turquoise shows significant anti-correlation with chemoresistance (-0.25 pval=0.008 ).
Module eigengene grey shows significant anti-correlation with chemoresistance (-0.2 pval=0.03 ).
Module eigengene yellow shows significant anti-correlation with chemoresistance (-0.19 pval=0.05 ).

### Gene relationship to trait and important modules:Gene ignificance and Module Membership  

We quantify associations of individual genes with our trait of interest (pCR) by defining Gene Significance GS as (the absolute value of) the correlation between the gene and the trait. For each module, we also define a quantitative measure of module membership MM as the correlation of the module eigengene and the gene expression profile. This allows us to quantify the similarity of all genes on the array to every module.  

```{r}
# Define variable weight containing the weight column of datTrait
ChemoRes = as.data.frame(datTrait$ChemoRes);
names(ChemoRes) = "ChemoRes"
# names (colors) of the modules
modNames = substring(names(MEs), 3)

geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");

geneTraitSignificance = as.data.frame(cor(datExpr, ChemoRes, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));

names(geneTraitSignificance) = paste("GS.", names(ChemoRes), sep="");
names(GSPvalue) = paste("p.GS.", names(ChemoRes), sep="");
```  

### Intramodular analysis: identfying genes with GS and MM  

Using the GS and MM measures, we can identify genes that have a high significance for pCR as well as high module membership in interesting modules.  

```{r}
MM_vs_GS <- function(module, modNames, moduleColors){
  column = match(module, modNames);
  moduleGenes = moduleColors==module;
  sizeGrWindow(7, 7);
  par(mfrow = c(1,1));
  verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                     abs(geneTraitSignificance[moduleGenes, 1]),
                     xlab = paste("Module Membership in", module, "module"),
                     ylab = "Gene significance for Chemoresistance",
                     main = paste("Module membership vs. gene significance\n"),
                     cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module, pch=19)
  abline(h=0.2, col = "red")
  abline(v=0.7, col = "red")
  return(NULL)
}

MM_vs_GS(module = "grey", modNames, moduleColors)
MM_vs_GS(module = "yellow", modNames, moduleColors)
MM_vs_GS(module = "turquoise", modNames, moduleColors)
```  

### Summary output of network analysis result  

We have found modules with high association with our trait of interest, and have identified their central players by the Module Membership measure. We now merge this statistical information with gene annotation and perform an enrichment analysis. After that we write out several files that summarizes the most important results.  

```{r message=FALSE, warning=FALSE, include=FALSE}
# Load gene annotation info from NS_CancerPath_C2535.rlf file

#annot <- readRlf("Data/RLFData/NS_CancerPath_C2535.rlf")
annot <- readRlf(paste0(DataDir, "/RLFData/NS_CancerPath_C2535.rlf")) # We can use DataDir defined at the beginning

get.probe <- function(complexid){
  parts = strsplit(complexid, split = "_")[[1]]
  paste("NM", parts[4], sep="_")
}

probes = unname(sapply(names(datExpr), get.probe))
probes2annot <-  match(probes, annot$Accession)
sum(is.na(probes2annot)) # number or probes without annotation. Should be 0

annot.geneid <- bitr(annot$GeneName, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
names(annot.geneid) <- c("GeneName", "ENTREZID")
annot <- merge(annot, annot.geneid, by="GeneName")
annot <- annot %>% dplyr::select(Accession, GeneName, ENTREZID, ProbeID)

## Define the universe backgroud for the enrichments
universe = unique(annot$ENTREZID) # Note that we are not using the universe in the enrichments

# Create the starting data frame
geneInfo0 = data.frame(Accession = probes,
                       geneSymbol = annot$GeneName[probes2annot],
                       entrezID = annot$ENTREZID[probes2annot],
                       moduleColor = moduleColors,
                       geneTraitSignificance,
                       GSPvalue)
# Order modules by their significance for weight
modOrder = order(-abs(cor(MEs, ChemoRes, use = "p")));
# Add module membership information in the chosen order
for (mod in 1:ncol(geneModuleMembership)){
  oldNames = names(geneInfo0)
  geneInfo0 = data.frame(geneInfo0, geneModuleMembership[, modOrder[mod]], 
                         MMPvalue[, modOrder[mod]]);
  names(geneInfo0) = c(oldNames, paste("MM.", modNames[modOrder[mod]], sep=""),
                       paste("p.MM.", modNames[modOrder[mod]], sep=""))
}

# Order the genes in the geneInfo variable first by module color, then by geneTraitSignificance
geneOrder = order(-abs(geneInfo0$GS.ChemoRes),geneInfo0$moduleColor);
geneInfo = geneInfo0[geneOrder, ]
write.csv(geneInfo, file = "Results/geneInfo.csv", quote = F,row.names = F)

## Write interesting modules gene info in a file
intModules <- c("grey","yellow","turquoise");
module_gene_info <- function(module, geneInfo){
  if(module=="grey"){
	  ChemoRes.genes <- geneInfo[which(geneInfo$moduleColor==module), c(1,2,3,4,5,6,9,10)]
	}
	if(module=="yellow"){
	  ChemoRes.genes <- geneInfo[which(geneInfo$moduleColor==module), c(1,2,3,4,5,6,11,12)]
	}
	if(module=="turquoise"){
	  ChemoRes.genes <- geneInfo[which(geneInfo$moduleColor==module), c(1,2,3,4,5,6,7,8)]
	}
	write.csv(ChemoRes.genes, file = paste("Results/",module,"-genesInfo.csv", sep=""), quote = F,row.names = F)
	return(NULL)
}
lapply(intModules, module_gene_info, geneInfo)
```  

## Interfacing with DIAMOnD  

```{r}
# Load saved data (datExpr and datTrait)
load("Data/RData/chemoresNAC_wgcna-01-dataInput.RData")
load("Data/RData/chemoresNAC_wgcna-02-networkConstruction.RData")
```  

Now, we have to download several high confidence human interactomes in order to run the DIAMOnD algorithm. With the interactome and the seed genes we can use the algorithm to expand the interesting modules and construct networks from them. We will use:  

* The DIAMOnD paper interactome provided by Ghiassian et al.(2015)[1].  
* STRINGdb human interactome with a score threshold > 0.7.  
* iRef human interactome  
* A metabolism centered interactome

### Proccessing the STRINGdb interactome  
We download the whole interactome of Homo sapiens from https://string-db.org/cgi/download.pl?sessionId=ED2PAnsg3O9r&species_text=Homo+sapiens. We need to process the data frame to apply a score filter and to map the ensembl protein ids to gene ids.  

```{r eval=FALSE, include=FALSE}
# Read file
string_interactome <- read.table(file = "Data/stringdb_9606.protein.links.v11.0.txt", header = TRUE)
string_interactome <- string_interactome %>% filter(combined_score > 700) %>% 
  tidyr::separate(col = protein1, into = c("organism1","N1"), sep = '\\.', remove = T) %>% 
  tidyr::separate(col = protein2, into = c("organism2","N2"), sep = '\\.', remove = T) %>%
  dplyr::select(N1, N2, combined_score)

ensembl = biomaRt::useMart("ensembl", dataset = "hsapiens_gene_ensembl")

# Dictionary ensembl protein id to gene id construction
interactome_ids <- unique(c(string_interactome$N1,string_interactome$N2))
dict <- getBM(attributes = c("ensembl_peptide_id","entrezgene"), 
      filters = "ensembl_peptide_id",
      values = dict,
      mart = ensembl)

# Mapping ensembl protein ids to entrez gene ids
ENSP_to_geneID <- function(ENSP_id, dict){
  if(ENSP_id %in% dict$ensembl_peptide_id){
    indx <- match(ENSP_id, dict$ensembl_peptide_id)
    return(dict$entrezgene[indx])
  }else{
    return(NA)
  }
}

string_interactome <- na.omit(string_interactome %>% 
  dplyr::mutate(gene_id1 = unlist(lapply(string_interactome$N1, ENSP_to_geneID, dict))
                                  , gene_id2 = unlist(lapply(string_interactome$N2, ENSP_to_geneID, dict))) %>%
  dplyr::select(gene_id1, gene_id2, combined_score))

# Save processed interactome in a file
write.table(string_interactome, row.names = F, col.names = T, file = "Data/processed_STRINGdb_Hs_interactome.tsv")
```

### Loading and processing iRef interactome  

```{r eval=FALSE, include=FALSE}
##irefindex_tab <- iRefR::get_irefindex(tax_id = "9606", iref_version = "13.0", data_folder = "Data")
load(file = "Data/9606.mitab.08122013.RData")

# creating an id conversion table to translate icROGs to entrez gene ids
id_conversion_table = create_id_conversion_table(irefindex_tab,
                                            data_folder = "Data",
                                            output_filename = "id_conversion_table_9606", 
                                            IDs_to_include = "entrezgene/locuslink")

load(file="Data/id_conversion_table_9606.RData")

# Mapping protein iROGs to gene ids
iROGid2geneid <- function(irogid, dict){
  if(irogid %in% dict$irogid){
    indx <- match(irogid, dict$irogid)
    return(dict$id_value[indx])
  }else{
    return(NA)
  }
}

# Apply the previous function, omit rows with missing values and keep gene ids and confidence.
iRef_interactome <- na.omit(irefindex_tab %>% 
                              dplyr::select(irogida, irogidb, confidence) %>%
                              dplyr::mutate(gene_id1 = as.numeric(unlist(lapply(irogida, 
                                                       iROGid2geneid, 
                                                       as.data.frame(id_conversion_table)))),
                                            gene_id2 = as.numeric(unlist(lapply(irogidb, 
                                                       iROGid2geneid, 
                                                       as.data.frame(id_conversion_table))))) %>% 
                              dplyr::select(gene_id1, gene_id2, confidence)) %>% 
                              dplyr::distinct();
# Filter the interactions on basis on confidence
## ¿?

# Save processed interactome in a file.
write.table(iRef_interactome, file = "Data/iRef_interactome.tsv", row.names = F, col.names = T)
```  

### Loading and processing the metabolism centered interactome  

We download the interactome from https://bioinfo.uth.edu/ccmGDB/download/2072genes_PathwayCommons.txt?csrt=5379671582968607161. This interactome has Pathway Commons format (see http://www.pathwaycommons.org/pc/sif_interaction_rules.do). We have to process it to keep only the INTERACTS_WITH relationships.  

```{r eval=FALSE, include=FALSE}
met.interactome <- na.omit(read.table(file = "Data/2072genes_PathwayCommons.txt", header = F, stringsAsFactors = F, fill = T, col.names = c('??','entity1','relationship_type','entity2'))) %>% dplyr::filter(relationship_type == 'interacts-with') %>% dplyr::select(entity1, entity2)

symbol_ids <- unique(c(met.interactome$entity1, met.interactome$entity2))
dict <- bitr(symbol_ids, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")

symbol_to_entrez <- function(symbol_id, dict){
  indx <- match(symbol_id, dict$SYMBOL)
  if(!is.na(indx)){
    return(dict$ENTREZID[indx])
  }else{
    return(NA)
  }
}

met.interactome <- na.omit(met.interactome %>% 
  mutate(N1 = as.numeric(unlist(lapply(entity1, symbol_to_entrez, dict))), 
         N2 = as.numeric(unlist(lapply(entity2, symbol_to_entrez, dict))) %>% 
  dplyr::select(N1, N2))

# Save processed interactome in a file.
write.table(met.interactome, file = "Data/metabolism-centered_interactome.tsv", row.names = F, col.names = T)
rm(symbol_ids)
rm(dict)
rm(met.interactome)
```  

Finally, we keep 684776 interactions.  

### Running DIAMOnD  

```{r}
# pybin : Path to python source
# network_path: Path to network edge list file
# seed_genes_path: Path to seed genes list file
# n: Number of DIAMOnD iterations
# alpha: Weight of the seeds, default value is set to 1
# diamond_path : DIAMOnD path
# outfilename : Output file name

run_DIAMOnD <- function(pybin, diamond_path, network_path, seed_genes_path, n, alpha = 1, outfilename){
  system(paste(pybin, diamond_path, network_path, seed_genes_path, n, alpha, outfilename, sep = " "))
}

```  

## WGCNA modules enrichment analysis  

### Loading data   

```{r}
# Load genes info
geneInfo <- read.csv(file = "Results/geneInfo.csv", header = T)

# Separate modules
grey_module <- geneInfo %>% 
  dplyr::filter(moduleColor == "grey" & abs(MM.grey) >= 0.70 & abs(GS.ChemoRes) >= 0.15) %>%
  dplyr::select(Accession, geneSymbol, entrezID, GS.ChemoRes, p.GS.ChemoRes, MM.grey, p.MM.grey)

yellow_module <- geneInfo %>% 
  dplyr::filter(moduleColor == "yellow" & abs(MM.yellow) >= 0.70 & abs(GS.ChemoRes) >= 0.15) %>%
  dplyr::select(Accession, geneSymbol, entrezID, GS.ChemoRes, p.GS.ChemoRes, MM.yellow, p.MM.yellow)

turquoise_module <- geneInfo %>% 
  dplyr::filter(moduleColor == "turquoise" & abs(MM.turquoise) >= 0.70 & abs(GS.ChemoRes) >= 0.15) %>%
  dplyr::select(Accession, geneSymbol, entrezID, GS.ChemoRes, p.GS.ChemoRes, MM.turquoise, p.MM.turquoise)
```  

### Modules enrichment analysis  

```{r}
source("Code/chemoRes_enrichment.R")

module_enrichment_analysis <- function(geneInfo, module, GS, MM){
  
  MM.module <- paste("MM.",module,sep="")
  
  res <- geneInfo %>% 
    dplyr::filter(moduleColor == module & !is.na(entrezID) & abs(GS.ChemoRes) >= GS & abs(geneInfo[,MM.module]) >= MM) %>%
    dplyr::select(Accession, geneSymbol, entrezID, GS.ChemoRes) %>% 
    dplyr::mutate(entrezID = as.character(entrezID))
     
  genes <- res$entrezID
  
  enrich_rs <- enrich_cp.chemoRes(genes, comparison = "ChemoRes")
  enrich_summary <- enrich_rs$summary %>% arrange(qvalue)
  enrich_summary <- convert_enriched_ids(enrich_summary, entrezsymbol = entrezsymbol) %>% filter(qvalue <=0.01) %>% arrange(-Count)
  
  write.csv(res, file = paste("Results/modules_enrichments/",module,"_module_genes.csv", sep="") )
  write.csv(enrich_summary, file = paste("Results/modules_enrichments/",module,"_module_enrichment.csv", sep=""))
}

module_enrichment_analysis(geneInfo, "grey", GS=0.1, MM = 0.2)
module_enrichment_analysis(geneInfo, "yellow", GS=0.1, MM=0.7)
module_enrichment_analysis(geneInfo, "turquoise", GS=0.1, MM=0.7)
```  

## Modules expansion with DIAMOnD  

```{r}
expand_module <- function(module, n, network_path){
  pybin <- "/anaconda3/bin/python"
  diamond_path <- "Code/DIAMOnD.py"
  interactome <- unlist(strsplit(network_path, "/|\\."))[2]
  
  aux.df <- read.csv(file = paste("Results/kk/",module,"_module_genes.csv", sep=""))
  genes <- as.numeric(aux.df$entrezID)
  seed_genes_path <- paste("Data/",module,"_seed_genes.txt", sep="")
  diamond_res_outfilename <- "Results/DIAMOnD_res.csv"
  
  # Write temporary module seed genes file
  write.table(genes ,file = seed_genes_path, row.names = F, col.names = F)
  
  run_DIAMOnD(pybin, diamond_path, network_path, seed_genes_path, n, outfilename = diamond_res_outfilename)
  
  aux.df <- read.csv(file = diamond_res_outfilename, header = T, sep = "\t")
  new_genes <- aux.df$DIAMOnD_node
  expanded_module_genes <- c(genes, new_genes)
  
  # Delete temporary module seed genes file
  system(paste("rm", seed_genes_path, sep=" "))
  # Delete temporary DIAMOnD result file
  system(paste("rm", diamond_res_outfilename, sep=" "))
  res.filename <- paste("Results/expanded_modules/",interactome,"_n=",n,"_expanded_",module,"_module.txt", sep = "")
  
  cat(paste("Expanded ",module, " module genes in ",res.filename, sep = ""))
  # Write module expanded genes file
  write.table(expanded_module_genes, file = res.filename, row.names = F, col.names = F)
}

# n = 300
expand_module("yellow", network_path = "Data/Ghiassian_interactome.tsv", n=300)
expand_module("grey", network_path = "Data/Ghiassian_interactome.tsv", n=300)
expand_module("turquoise", network_path = "Data/Ghiassian_interactome.tsv", n=300)

expand_module("yellow", network_path = "Data/iRef_interactome.tsv", n=300)
expand_module("grey", network_path = "Data/iRef_interactome.tsv", n=300)
expand_module("turquoise", network_path = "Data/iRef_interactome.tsv", n=300)

expand_module("yellow", network_path = "Data/STRINGdb_interactome.tsv", n=300)
expand_module("grey", network_path = "Data/STRINGdb_interactome.tsv", n=300)
expand_module("turquoise", network_path = "Data/STRINGdb_interactome.tsv", n=300)

expand_module("yellow", network_path = "Data/metabolism-centered_interactome.tsv", n=300)
expand_module("grey", network_path = "Data/metabolism-centered_interactome.tsv", n=300)
expand_module("turquoise", network_path = "Data/metabolism-centered_interactome.tsv", n=300)

## n = 200
expand_module("yellow", network_path = "Data/Ghiassian_interactome.tsv", n=200)
expand_module("grey", network_path = "Data/Ghiassian_interactome.tsv", n=200)
expand_module("turquoise", network_path = "Data/Ghiassian_interactome.tsv", n=200)

expand_module("yellow", network_path = "Data/iRef_interactome.tsv", n=200)
expand_module("grey", network_path = "Data/iRef_interactome.tsv", n=200)
expand_module("turquoise", network_path = "Data/iRef_interactome.tsv", n=200)

expand_module("yellow", network_path = "Data/STRINGdb_interactome.tsv", n=200)
expand_module("grey", network_path = "Data/STRINGdb_interactome.tsv", n=200)
expand_module("turquoise", network_path = "Data/STRINGdb_interactome.tsv", n=200)

expand_module("yellow", network_path = "Data/metabolism-centered_interactome.tsv", n=200)
expand_module("grey", network_path = "Data/metabolism-centered_interactome.tsv", n=200)
expand_module("turquoise", network_path = "Data/metabolism-centered_interactome.tsv", n=200)

## n = 100
expand_module("yellow", network_path = "Data/Ghiassian_interactome.tsv", n=100)
expand_module("grey", network_path = "Data/Ghiassian_interactome.tsv", n=100)
expand_module("turquoise", network_path = "Data/Ghiassian_interactome.tsv", n=100)

expand_module("yellow", network_path = "Data/iRef_interactome.tsv", n=100)
expand_module("grey", network_path = "Data/iRef_interactome.tsv", n=100)
expand_module("turquoise", network_path = "Data/iRef_interactome.tsv", n=100)

expand_module("yellow", network_path = "Data/STRINGdb_interactome.tsv", n=100)
expand_module("grey", network_path = "Data/STRINGdb_interactome.tsv", n=100)
expand_module("turquoise", network_path = "Data/STRINGdb_interactome.tsv", n=100)

expand_module("yellow", network_path = "Data/metabolism-centered_interactome.tsv", n=100)
expand_module("grey", network_path = "Data/metabolism-centered_interactome.tsv", n=100)
expand_module("turquoise", network_path = "Data/metabolism-centered_interactome.tsv", n=100)
```  

## Expanded modules enrichment  

With DIAMOnD we have obtained networks built on basis on modules initial seed genes and the diferent interactomes. Since we had 3 interactomes and 3 modules, now we have 9 networks. The next step is to evaluate the networks by performing a functional enrichment.  

```{r eval=FALSE, include=FALSE}
# file.remove(list.files(path = "Results/expanded_modules_enrichments", full.names = T))
# expanded_modules <- list.files(path = "Results/expanded_modules", full.names = T)

expanded_module_enrichment <- function(genes_file_path){
  outfilename <- paste("Results/expanded_modules_enrichments/",unlist(strsplit(genes_file_path, '/|\\.'))[3], "_enrichment.csv",sep = "")
  genes <- readLines(con = genes_file_path)
  
  cat(paste(genes_file_path, "enrichment...", '\n'))
  
  enrich_rs <- enrich_cp.chemoRes(genes, comparison = "ChemoRes")
  enrich_summary <- enrich_rs$summary %>% arrange(qvalue)
  enrich_summary <- convert_enriched_ids(enrich_summary, entrezsymbol = entrezsymbol) %>% filter(qvalue <=0.01) %>% arrange(-Count)
  
  write.csv(enrich_summary, file = outfilename)
  
  cat(paste(genes_file_path, "enrichment DONE", '\n'))
  cat(paste("Results in", outfilename, '\n'))
  return("OK")
}

lapply(expanded_modules, expanded_module_enrichment)
```  

## Network construction from expanded modules  

```{r}
# file.remove(list.files(path = "Results/expanded_modules_networks", full.names = T))
expanded_modules <- list.files(path = "Results/expanded_modules", full.names = T)

entrez2symbol <- function(entrez, dict){
  indx <- match(entrez, dict$ENTREZID)
  if(!is.na(indx)){
    return(dict$SYMBOL[indx])
  }else{
    return(NA)
  }
}

network_building <- function(genes_file_path){
  expanded_module <- unlist(strsplit(genes_file_path, '/|\\.'))[3]
  outfilename <- paste("Results/expanded_modules_networks/",expanded_module, "_network.csv",sep = "")
  interactome_name <- paste(unlist(strsplit(expanded_module, '_'))[1], 
                            unlist(strsplit(expanded_module, '_'))[2],
                            sep='_')
  
  interactome_file_path <- paste("Data/",interactome_name,".tsv",sep="")
  
  genes <- as.numeric(readLines(con = genes_file_path))
  dict <- bitr(genes, fromType="ENTREZID", toType="SYMBOL", OrgDb="org.Hs.eg.db")
  
  if(interactome_file_path == 'Data/STRINGdb_interactome.tsv'){
    interactome <- read.table(file = interactome_file_path, header = F)[1:3]
    colnames(interactome) <- c('gene_id1', 'gene_id2','score')
    
    network <- interactome %>% 
      dplyr::filter(gene_id1 %in% genes & gene_id2 %in% genes)
  }else{
    interactome <- read.table(file = interactome_file_path, header = F)[1:2]
    colnames(interactome) <- c('gene_id1', 'gene_id2')
  
    network <- interactome %>% 
      dplyr::filter(gene_id1 %in% genes & gene_id2 %in% genes)
  }
  
  gene_symbol1 <- unlist(lapply(network$gene_id1, entrez2symbol, dict))
  gene_symbol2 <- unlist(lapply(network$gene_id2, entrez2symbol, dict))
  network <- network %>%
    dplyr::mutate(gene_symbol1 = gene_symbol1, 
                  gene_symbol2 = gene_symbol2)
  
  write.csv(network, file = outfilename)
}

lapply(expanded_modules, network_building)
```

## References  
1. Ghiassian SD, Menche J, Barabási A-L (2015) A DIseAse MOdule Detection (DIAMOnD) Algorithm Derived from a Systematic Analysis of Connectivity Patterns of Disease Proteins in the Human Interactome. PLoS Comput Biol 11(4): e1004120. https://doi.org/10.1371/journal.pcbi.1004120

